
---
title: "GroupComparison Package Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GroupComparison Package Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The `GroupComparison` package provides a set of functions to visualize group comparisons using violin plots. These functions allow for flexibility in customizing the visual elements such as confidence intervals, p-values, medians, outliers, and more.

This vignette demonstrates the usage of the three key functions in the package:

1. `create_violin_plot()`: Create violin plots with flexible options for visualizing medians, outliers, boxplots, and p-values.
2. `violin_plots_with_comparison_group()`: Create violin plots with a designated comparison group, including p-values for pairwise comparisons.
3. `split_violin_plot()`: Create split violin plots that divide the violin shape based on a split criterion (e.g., a binary condition).

## 1. Violin Plot with Custom Elements

The `create_violin_plot()` function is used to create violin plots with additional elements like boxplots, medians, outliers, and p-values for group comparisons.

### Example

```{r violin-plot, message=FALSE, warning=FALSE}
library(GroupComparison)
library(ggplot2)

# Example data
data <- data.frame(
  group = rep(c("A", "B", "C"), each = 100),
  value = c(rnorm(100, mean = 0), rnorm(100, mean = 1), rnorm(100, mean = 2))
)

# Custom group colors
group_colors <- c('A' = "#8DA0CB80", 'B' = "#66C2A580", 'C' = "#E41A1C80")

# Create a violin plot
p <- create_violin_plot(
  data = data,
  group_column = "group",
  value_column = "value",
  group_colors = group_colors,
  p_value = TRUE,
  p_value_format = "asterisk",
  y_lab = bquote(bold(Delta ~ Log(activity))),
  breaks = c(-3, 0, 1, 2, 3),
  limits = c(-3, 3)
)

# Plot the violin plot
print(p)
```

## 2. Violin Plots with a Comparison Group

The `violin_plots_with_comparison_group()` function creates violin plots for various groups, with a designated comparison group highlighted. It also provides p-values for pairwise comparisons with the comparison group.

### Example

```{r violin-plot-comparison, message=FALSE, warning=FALSE}
# Example data
data <- data.frame(
  group = rep(c("A", "B", "C"), each = 100),
  value = c(rnorm(100, mean = 0), rnorm(100, mean = 1), rnorm(100, mean = 2))
)

# Create a violin plot with a comparison group
p_comparison <- violin_plots_with_comparison_group(
  data = data,
  group_column = "group",
  value_column = "value",
  comparison_group = "A",
  p_value = TRUE,
  p_value_format = "asterisk",
  y_lab = bquote(bold(Delta ~ Log(activity))),
  breaks = c(-3, 0, 1, 2, 3),
  limits = c(-3, 3)
)

# Plot the violin plot with comparison group
print(p_comparison)
```

## 3. Split Violin Plot

The `split_violin_plot()` function creates split violin plots, where the violin is divided into two halves based on a split criterion. This is useful for comparing two subgroups within each group.

### Example

```{r split-violin-plot, message=FALSE, warning=FALSE}
# Example data with a split criterion
data <- data.frame(
  group = rep(c("A", "B", "C"), each = 100),
  value = c(rnorm(100, mean = 0), rnorm(100, mean = 1), rnorm(100, mean = 2)),
  split_criteria = sample(c(TRUE, FALSE), 300, replace = TRUE)
)

# Colors for the split halves
colors <- c("TRUE" = "#FF9999", "FALSE" = "#99CCFF")

# Labels for the legend
labels <- c("TRUE" = "First Group", "FALSE" = "Second Group")

# Create a split violin plot
p_split <- split_violin_plot(
  data = data,
  group_column = "group",
  value_column = "value",
  split_column = "split_criteria",
  colors = colors,
  labels = labels,
  x_lab = "Groups",
  y_lab = bquote(bold(Delta ~ Log(activity))),
  outliers = TRUE,
  CI = TRUE,
  median = TRUE,
  n_obs = TRUE
)

# Plot the split violin plot
print(p_split)
```

# Conclusion

The `GroupComparison` package provides a flexible and powerful way to create customized violin plots for comparing multiple groups. The package's functions allow for extensive customization, including p-values, confidence intervals, medians, outliers, and more.
